---
title: "Continuous Glucose Monitoring vs. Blood Glucose Monitoring in Non-Insulin-Treated Type 2 Diabetes"
subtitle: "A reproducible meta-analysis of HbA1c outcomes"
author: "Said Adjao"
date: today
format:
  html:
    toc: true                 # add a table of contents and numbered headings to make it read like a real report
    number-sections: true     # add a table of contents and numbered headings to make it read like a real report
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup

library(tidyverse)
library(meta)

cgm <- tibble(
  study   = c("Wada 2020","Aronson 2023 (IMMEDIATE)","Lau 2024","Ssemmondo 2025",
                    "Cox 2020","Moon 2023","Price 2021 (COMMITTED)","Allen 2008"),
  device  = c("isCGM","isCGM","isCGM","isCGM",
             "rtCGM","rtCGM","rtCGM","Professional"),
  te     = c(-0.29, -0.30, -0.65, -0.73, -1.11, -0.60, -0.30, -0.84),
  se     = c(0.127, 0.152, 0.245, 0.513, 0.606, 0.254, 0.904, 0.429),
  source = c("Back-calculated","Back-calculated","Back-calculated","Primary paper",
             "Primary paper","Back-calculated","Back-calculated","Back-calculated")
)


m <- metagen(
  TE = cgm$te, 
  seTE = cgm$se,
  studlab = cgm$study,
  sm = "MD",
  common = FALSE,
  random  = TRUE
)
```

## Summary

This report synthesizes randomized evidence on continuous glucose
monitoring (CGM) versus blood glucose monitoring (BGM) for glycaemic
control in adults with type 2 diabetes **not treated with insulin** — a
population for which CGM coverage is frequently restricted. Pooling
`r m$k` trials, CGM was associated with a mean HbA1c reduction of
**`r round(m$TE.random, 2)`%** (95% CI `r round(m$lower.random, 2)` to
`r round(m$upper.random, 2)`), with no detectable between-study
heterogeneity (I² = `r round(m$I2 * 100, 1)`%). The estimate is
consistent with a recently published meta-analysis, supporting a modest
but reliable benefit in this population.

## Background

Most clinical-trial evidence and payer coverage for CGM in type 2
diabetes derives from insulin-treated populations. Whether the benefit
extends to non-insulin-treated patients — the larger group, and the one
typically excluded from CGM reimbursement — is a distinct question with
direct cost and coverage implications. This analysis synthesizes the
randomized evidence specific to that population.

## Methods

### Eligibility criteria

Randomized controlled trials were eligible if they enrolled adults with
type 2 diabetes treated **exclusively with non-insulin therapies**,
compared any CGM modality (intermittently scanned, real-time, or
professional) against BGM or usual care, and reported HbA1c as an
outcome. Trials enrolling any insulin-treated participants were excluded.

### Information sources and study selection

The trial set was harvested from a 2025 systematic review by Aronson and
colleagues that applied the same non-insulin-only inclusion criterion,
providing a transparent and reproducible starting point. Eight trials
met criteria.

### Data extraction

HbA1c effect estimates were obtained by two routes. For two trials, the
between-group mean difference and its dispersion were extracted directly
from the primary publications. For the remaining six, the standard error
was back-calculated from the reported mean difference and its p-value or
confidence interval — a standard technique where standard errors are not
printed directly. The extraction route for each trial is shown in the
results table. One trial reporting HbA1c in IFCC units (mmol/mol) was
converted to DCCT/NGSP units (%) prior to pooling.

### Statistical analysis

Effects were pooled using an inverse-variance random-effects model
(mean difference scale), implemented with the `meta` package in R.
Random effects were chosen a priori because the trials differed in CGM
modality, follow-up duration, and co-interventions, making a single
common true effect implausible. Heterogeneity was quantified with I².

## Results

```{r}
#| label: study-table

cgm %>%
  mutate(
    MD = round(te, 2),
    `95% CI` = paste0("(", round(te - 1.96 * se, 2), ", ",
                      round(te + 1.96 * se, 2), ")")
  ) %>%
  select(Study = study, Device = device, MD, `95% CI`,
         `Effect source` = source) %>%
  knitr::kable(caption = "HbA1c mean difference (CGM − BGM), % points. Negative favors CGM.")  # knitr::kable() formats the data as clean table with a caption, instead of raw console output
```

```{r}
#| label: forest-plot
#| fig-width: 9
#| fig-height: 5

forest(m,
       leftcols = c("studlab"),
       label.left = "Favors CGM",
       label.right = "Favors BGM")
```

Pooling all `r m$k` trials, CGM was associated with a significantly
greater HbA1c reduction than BGM (mean difference
**`r round(m$TE.random, 2)`%**; 95% CI `r round(m$lower.random, 2)` to
`r round(m$upper.random, 2)`). Between-study heterogeneity was absent
(I² = `r round(m$I2 * 100, 1)`%), though this should be interpreted
cautiously given the small number of small trials.

### Validation

This pooled estimate is consistent with the published benchmark
(−0.37%; 95% CI −0.49 to −0.24), reproducing a peer-reviewed result from
the primary data. Minor differences are attributable to the standard-error
derivation and missing-data imputation approaches.

## Limitations

Several limitations temper confidence in the pooled estimate. The
evidence base is small (`r m$k` trials, most with modest sample sizes),
limiting the precision of the heterogeneity estimate; an I² of 0% in
this setting reflects absence of *detected* heterogeneity rather than
strong evidence of consistency. All trials were open-label, raising the
possibility of performance and detection bias. Several interventions
bundled CGM with structured education or lifestyle programmes, so the
pooled effect reflects CGM-plus-support rather than the device in
isolation. Finally, six of eight effect estimates were back-calculated;
these should be verified against the published forest plot before any
formal use.

## References

1. Aronson R, et al. Continuous glucose monitoring in noninsulin-treated type 2 diabetes. *Diabetes Obes Metab.* 2025;27(11):6220–6242.
2. Aronson R, et al. IMMEDIATE trial. *Diabetes Obes Metab.* 2023;25(4):1024–1031.
3. Lau D, Manca DP, Singh P, et al. The effectiveness of continuous glucose monitoring with remote telemonitoring-enabled virtual educator visits in adults with non-insulin dependent type 2 diabetes: a randomized trial. *Diabetes Res Clin Pract.* 2024.
4. Ssemmondo E, Shah N, Newham M, et al. Effect of introduction of intermittently scanned continuous glucose monitoring on glycaemic control in individuals with type 2 diabetes treated with non-insulin therapies. *Diabetes Obes Metab.* 2025;27(3):1226–1232.
5. Cox DJ, Banton T, Moncrief M, et al. Minimizing glucose excursions (GEM) with continuous glucose monitoring in type 2 diabetes: a randomized clinical trial. *J Endocr Soc.* 2020;4(11):bvaa118.
6. Moon SJ, Kim KS, Lee WJ, et al. Efficacy of intermittent short-term use of a real-time continuous glucose monitoring system in non-insulin-treated patients with type 2 diabetes: a randomized controlled trial. *Diabetes Obes Metab.* 2023;25(1):110–120.
7. Price DA, Deng Q, Kipnes M, Beck SE. Episodic real-time CGM use in adults with type 2 diabetes (COMMITTED): results of a pilot randomized controlled trial. *Diabetes Ther.* 2021;12(7):2089–2099.
8. Allen NA, Fain JA, Braun B, Chipkin SR. Continuous glucose monitoring counseling improves physical activity behaviors of individuals with type 2 diabetes: a randomized clinical trial. *Diabetes Res Clin Pract.* 2008;80:371–379.
9. Wada E, Onoue T, Kobayashi T, et al. Flash glucose monitoring helps achieve better glycemic control than conventional self-monitoring of blood glucose in non-insulin-treated type 2 diabetes: a randomized controlled trial. *BMJ Open Diabetes Res Care.* 2020;8(1):e001115.




